---
title: "A mini-lab using base R and packages of dplyr and ggplot2"
author: "Jimmy Fang"
date: "10/12/2020"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Introduction

11921 users in a Chinese social Q&A community were tracked for a 20-week period, and key variables of users were collected. They include name, gender, count of answers, count of questions and count of followers, etc. In this lab, I will do some descriptive analysis (including drawing graphs) on this data set by using base R and packages of dplyr and ggplot2.

## Check packages and import dataset
```{r dataset}
library(dplyr)
library(pander)
library(ggplot2)
library(ggthemes)
library(gridExtra)

setwd("/Users/jimmyfang/Desktop/R Mini-lab Oct 12")
grassroots <- read.csv("grassroots.csv")

# define date
grassroots <- grassroots %>%
  mutate(date = substr(date,1,10),
         date = as.Date(date,"%Y-%m-%d"))
```

## Total counts by date
```{r total counts}
# base
attach(grassroots)
aggregate(cbind(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count), 
          by=list(Date = date), FUN=sum)
detach(grassroots)

# dplyr
grassroots %>%
  group_by(date) %>%
  select(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count) %>%
  summarise_each(funs = sum) %>%
  pander()
```

## Mean counts by date
```{r mean counts}
# base
attach(grassroots)
aggregate(cbind(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count), 
          by=list(Date = date), FUN=mean)
detach(grassroots)

# dplyr
grassroots %>%
  group_by(date) %>%
  select(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count) %>%
  summarise_each(funs = mean) %>%
  pander()
```

## Standard deviation of counts by date
```{r std counts}
# base
attach(grassroots)
aggregate(cbind(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count), 
          by=list(Date = date), FUN=sd)
detach(grassroots)

# dplyr
grassroots %>%
  group_by(date) %>%
  select(answer_count, articles_count, question_count, favorite_count, favorited_count, 
                follower_count, following_count,following_favlists_count, following_question_count, 
                following_topic_count, thanked_count, voteup_count) %>%
  summarise_each(funs = sd) %>%
  pander()
```

## Finding out the users with largest followers
```{r users with largest followers}
# base
attach(grassroots)
grassroots[order(date,-follower_count),][seq(1,238420,11921),c(16,1:2,19,4:6,8:11,22:23,26:27)]
detach(grassroots)

# dplyr
grassroots %>%
  arrange(date,desc(follower_count)) %>%
  slice(seq(1,238420,11921)) %>%
  select(c(16,1:2,19,4:6,8:11,22:23,26:27)) %>%
  pander()
```

## Finding out users with largest incremental followers
```{r users with largest incremental followers}
# base
grassroots1 <- grassroots[order(grassroots$date,
                                grassroots$follower_count),][,c(16,1:2,19,4:6,8:11,22:23,26:27)]
grassroots1$follower_diff <- c(NA,diff(grassroots1$follower_count))
grassroots1[seq(1,238420,11921),16] <- NA

attach(grassroots1)
grassroots1[order(date,-follower_diff),][seq(1,238420,11921),]
detach(grassroots1)

# dplyr
grassroots1 %>%
  arrange(date, desc(follower_diff)) %>%
  slice(seq(1,238420,11921)) %>%
  pander()

remove(grassroots1) # drop grassroot1
```

## ScatterPlot between answer count and voteup count in the 1st week
```{r scatterplot}
# base
plot(grassroots[c(1:11921),c(1,26)],,col="blue",pch=16,cex=0.5, main="Scatterplot")
abline(lm(voteup_count~answer_count, data=grassroots), col = "red", lwd = 3)

# ggplot2
ggplot(grassroots[c(1:11921),],aes(x=answer_count,y=voteup_count)) +
      geom_point(col="blue") +
      geom_smooth(method = 'lm',se=FALSE,col = "red") +
      labs(title = "ScatterPlot",x = "Height in Inches", y="Salary in Dollars") +
      theme_gdocs()+
      theme(plot.title = element_text(hjust = 0.5),plot.subtitle = element_text(hjust = 0.5))
```

## Boxplots of answer count by gender in the 1st week
```{r boxplot}
# base
grassroots$gender[grassroots$gender==1] <- "male"
grassroots$gender[grassroots$gender==0] <- "female"
grassroots$gender[grassroots$gender==-1] <- "unknown"
boxplot(answer_count ~ gender, data = grassroots[c(1:11921),], main = "Answer Count by Gender")
by(grassroots[c(1:11921),1],grassroots[c(1:11921),12], summary) # check five-number summary

# ggplot2
ggplot(grassroots[c(1:11921),], aes(x = gender, y = answer_count)) + 
       stat_boxplot(geom = "errorbar", width = 0.75, coef = 1.5) +
      geom_boxplot() +
      labs(title="Answer Count by Gender",
        x ="gender", y = "answer count") +
      theme_few()+
      theme(plot.title = element_text(hjust = 0.5))
```

## Histogram of answer count and follower count in the 1st week
```{r histogram}
# base
par(mfrow=c(2,1))
hist(grassroots[c(1:11921),1], breaks = seq(0,1240,by=10), freq = F,
     col = "green", xlim = c(0,100), main = "Histogram of Answer Count", xlab = "answer count")
hist(grassroots[c(1:11921),6], breaks = seq(0,100,by=10), freq = F,
     col = "orange", xlim = c(0,100), main = "Histogram of Follower Count", xlab = "follower count")
par(mfrow=c(1,1))

# ggplot2
Hist_answer <- ggplot(grassroots[c(1:11921),],aes(answer_count)) +
      geom_histogram(aes(y = ..density..),breaks = seq(0,1240,by=10),
                     fill = "green",color="black") +
      labs(title="Histogram",x ="answer count") +
      theme_classic()+
      theme(plot.title = element_text(hjust = 0.5, face="bold"))

Hist_follower <- ggplot(grassroots[c(1:11921),],aes(follower_count)) +
      geom_histogram(aes(y = ..density..),breaks = seq(0,100,by=10),
                     fill = "orange",color="black") +
      labs(title="Histogram",x ="follower count") +
      theme_classic()+
      theme(plot.title = element_text(hjust = 0.5, face="bold"))

grid.arrange(Hist_answer,Hist_follower)

remove(Hist_answer,Hist_follower) # drop Hist_answer and Hist_follower
```

## The trend of follower count of user 'hou-cheng-yao'
```{r trend}
# base
hcy_trend <- grassroots[which(grassroots$url_token == "hou-cheng-yao"),]
plot(hcy_trend$date, hcy_trend$follower_count, type = "b", col = "blue",
     main = "Trend of Follower Count", sub = "hou-cheng-yao", 
     xlab = "date", ylab = "follower count")

# ggplot2
ggplot(hcy_trend,aes(x=date,y=follower_count)) +
      geom_line(color = "blue") +
      labs(title = "Trend of Follower Count", subtitle = "hou-cheng-yao",
           x = "date", y="follower count") +
      theme_gdocs()+
      theme(plot.title = element_text(hjust = 0.5),plot.subtitle = element_text(hjust = 0.5))

remove(hcy_trend) # drop hcy_trend
```

## Conclusion
In this lab, I look into the dataset from many perspectives by using base R and packages of dplyr or ggplot2. The main objective is to display how to code with those packages, thus neglecting the implication of results. Interested readers can work on them. All related resources on those packages could be found at https://www.r-project.org. I am particularly grateful to the authors who created those two smart packages.

